# Adding panel settings {#add-panel-setting}

## Create a slot

We define slot names and the type of value that they expect.

```{r}
collated <- character(0)
.rownamesShown <- "ShowRowNames"
.colnamesShown <- "ShowColumnNames"
collated[.rownamesShown] <- "logical"
collated[.colnamesShown] <- "logical"
# ... other slots ...
#' @export
setClass("ComplexHeatmapPlot", contains="Panel", slots=collated)
```

## Define new constants (optional)

If the panel setting involves a set of predefined choices - as opposed to dynamic sets choices that depend on the dataset such as feature names -, we may need to declare those predefined choices as constants.

In this case, both settings simply toggle between `TRUE` and `FALSE`, so there is no need for new constants.

## Initialize setting with a default value

In the `initialize` function, we set both settings to `TRUE`.
This indicates that both row and column names should be displayed by default.

```{r}
#' @export
#' @importFrom methods callNextMethod
setMethod("initialize", "ComplexHeatmapPlot", function(.Object, ...) {
    args <- list(...)

    # ... other slots ...
    args <- .empty_default(args, .rownamesShown, TRUE)
    args <- .empty_default(args, .colnamesShown, TRUE)

    do.call(callNextMethod, c(list(.Object), args))
})
```

