```{r, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, error = FALSE, cache = FALSE)
suppressPackageStartupMessages(require(BiocStyle))
options(digits = 4)
```

# Annotated gene list

## Overview

When given a gene list, we often need to look up the function of the top genes in a search engine.
This typically involves copy-pasting the gene name or ID into the search box and pressing Enter, which is a pain.
Instead, we can automate this process in `r BiocStyle::Biocpkg("iSEE")` by creating an **annotated gene table**.
We demonstrate by showing how we can dynamically look up annotation for each gene in the `rowData` of a `SummarizedExperiment`.

## Class basics 

First, we define the basics of our new `Panel` class.
Our new class will be showing the gene-level metadata, so we inherit from the `RowStatTable` class that does exactly this.
We add some slots specifying which column of the table contains our gene IDs, the type of ID and the organism database to use.

```{r, echo=FALSE}
library(iSEE)
library(S4Vectors)
setClass("GeneAnnoTable", contains="RowStatTable",
    slots=c(
        IDColumn="character_OR_NULL", 
        IDType="character",
        Organism="character",
        AnnoBoxOpen="logical"
    )
)
```

We specialize the validity method to check that the `IDColumn` is either a string or `NULL`;
if the latter, we assume that the ID is stored in the row name.
We also add some cursory checks for the other parameters.

```{r, results='hide'}
allowable <- c("ENSEMBL", "SYMBOL", "ENTREZID")
setValidity2("GeneAnnoTable", function(object) {
    msg <- character(0)

    if (!is.null(val <- object[["IDColumn"]]) && (length(val)!=1L || is.na(val))) {
        msg <- c(msg, "'IDColumn must be NULL or a string")
    }

    if (!isSingleString(orgdb <- object[["Organism"]])) {
        msg <- c(msg, sprintf("'Organism' should be a single string", orgdb))
    }

    if (!isSingleString(type <- object[["IDType"]]) || !type %in% allowable) {
        msg <- c(msg, "'IDType' should be 'ENSEMBL', 'SYMBOL' or 'ENTREZID'")
    }

    if (length(open <- object[["AnnoBoxOpen"]])!=1L || is.na(open)) {
        msg <- c(msg, "'AnnoBoxOpen' should be a non-missing logical scalar")
    }

    if (length(msg)) {
        return(msg)
    }
    TRUE
})
```

We then specialize the initialize method to set reasonable defaults for these parameters.

```{r}
setMethod("initialize", "GeneAnnoTable", function(.Object, IDColumn=NULL, 
    Organism="org.Mm.eg.db", IDType="SYMBOL", AnnoBoxOpen=FALSE, ...)
{
    callNextMethod(.Object, IDColumn=IDColumn, IDType=IDType,
        Organism=Organism, AnnoBoxOpen=AnnoBoxOpen, ...)
})
```

## Setting up the interface

We define the full name and desired default color for display purposes:

```{r}
setMethod(".fullName", "GeneAnnoTable", function(x) "Annotated gene table")

setMethod(".panelColor", "GeneAnnoTable", function(x) "#AA1122")
```

We want to add another UI element for showing the gene-level annotation.
This is achieved by specializing the `.defineOutput()` method as shown below;
note the prefixing by the panel name to ensure that output element IDs from different panels are unique.

```{r}
setMethod(".defineOutput", "GeneAnnoTable", function(x, ...) {
    panel_name <- .getEncodedName(x)
    tagList(
        callNextMethod(), # Re-using RowStatTable's definition.
        uiOutput(paste0(panel_name, "_annotation")),
        hr()
    )
})
```

We also set up interface elements for changing the annotation parameters.
We will put these elements in a separate "Annotation parameters" collapsible box,
which is initialized in an opened or closed state depending on the `AnnoBoxOpen` slot.

```{r}
setMethod(".defineInterface", "GeneAnnoTable", function(x, se, select_info) {
    panel_name <- .getEncodedName(x)
    c(
        list(
            collapseBox(
                paste0(panel_name, "_AnnoBoxOpen"),
                title="Annotation parameters",
                open=x[["AnnoBoxOpen"]],
                selectInput(paste0(panel_name, "_IDColumn"),
                    label="ID-containing column:",
                    choices=colnames(rowData(se)), 
                    selected=x[["IDColumn"]]
                ),
                selectInput(paste0(panel_name, "_IDType"),
                    label="ID type:",
                    choices=allowable,
                    selected=x[["IDType"]]
                ),
                selectInput(paste0(panel_name, "_Organism"),
                    label="Organism",
                    choices=c("org.Hs.eg.db", "org.Mm.eg.db"),
                    selected=x[["Organism"]]
                )
            )
        ),
        callNextMethod()
    ) 
})
```

## Creating the observers

We specialize `.createObservers` to define some observers to respond to changes in our new interface elements.
Note the use of `callNextMethod()` to ensure that observers of the parent class are also created.

```{r}
setMethod(".createObservers", "GeneAnnoTable",
    function(x, se, input, session, pObjects, rObjects)
{
    callNextMethod()

    plot_name <- .getEncodedName(x)

    .createUnprotectedParameterObservers(plot_name,
        fields=c("IDColumn", "Organism", "IDType"), 
        input=input, pObjects=pObjects, rObjects=rObjects)
})
```

We need to set up a rendering expression for the annotation element that responds to the selected gene.
By using `.trackSingleSelection()`, we ensure that this UI element updates in response to changes in the table selection.
We add a series of protective measures to avoid the application crashing due to missing organism packages or unmatched IDs.

```{r}
setMethod(".renderOutput", "GeneAnnoTable", function(x, se, ..., output, pObjects, rObjects) {
    callNextMethod() # Re-using RowStatTable's output rendering.

    panel_name <- .getEncodedName(x)
    output[[paste0(panel_name, "_annotation")]] <- renderUI({
        .trackSingleSelection(panel_name, rObjects)
        instance <- pObjects$memory[[panel_name]]

        rowdata_col <- instance[["IDColumn"]]
        selectedGene <- instance[["Selected"]]
        if (!is.null(rowdata_col)) {
            selectedGene <- rowData(se)[selectedGene,rowdata_col]
        }

        keytype <- instance[["IDType"]]
        selgene_entrez <- NA
        if (keytype!="ENTREZID") {
            ORG <- instance[["Organism"]]
            if (require(ORG, character.only=TRUE, quietly=TRUE)) {
                orgdb <- get(ORG)
                selgene_entrez <- try(mapIds(orgdb, selectedGene, "ENTREZID", keytype), 
                    silent=TRUE)
            }
        } else {
            selgene_entrez <- selectedGene
        }

        if (is.na(selgene_entrez) || is(selgene_entrez, "try-error")) {
            return(NULL)
        }

        fullinfo <- rentrez::entrez_summary("gene", selgene_entrez)
        link_pubmed <- paste0('<a href="http://www.ncbi.nlm.nih.gov/gene/?term=',
            selgene_entrez,
            '" target="_blank">Click here to see more at the NCBI database</a>')

        mycontent <- paste0("<b>",fullinfo$name, "</b><br/><br/>",
            fullinfo$description,"<br/><br/>",
            ifelse(fullinfo$summary == "","",paste0(fullinfo$summary, "<br/><br/>")),
            link_pubmed)

        HTML(mycontent)
    })
})
```

Observant readers will note that the body of the rendering expression uses `instance` rather than `x`.
This is intentional as it ensures that we are using the parameter settings from the current state of the app.
If we used `x`, we would always be using the parameters from the initial state of the app, which is not what we want.

## In action

Let's put our new panel to the test using the `sce` object, preprocessed in a [previous chapter](#developing).

```{r, include=FALSE}
sce <- readRDS("sce.rds")
```

Setting up the iSEE instance is as easy as:

```{r}
gat <- new("GeneAnnoTable", PanelWidth=8L)
app <- iSEE(sce, initial=list(gat))
```

Clicking on any row will bring up the Entrez annotation (if available) for that feature.
It is probably best to click on some well-annotated genes as the set of RIKEN transcripts at the front don't have much annotation.
