# Panel classes {#panels}

```{r, echo=FALSE, results="hide"}
knitr::opts_chunk$set(error=FALSE, warning=FALSE, message=FALSE)
suppressPackageStartupMessages(library(BiocStyle))
suppressPackageStartupMessages(library(iSEE))
class_link <- function(x) sprintf("[`%s`](https://isee.github.io/iSEE/reference/%s-class.html)", x, x)
```

## Overview

This chapter provides a list of all of the classes that are implemented by the core `r Biocpkg("iSEE")` package.
Each class comes with its specialized implementations of methods for various generics described in Chapter \@ref(api).
Thus, it is often possible for developers to inherit from one of these classes to get most of the methods implemented "for free".
The classes themselves are either virtual or concrete; the latter can be created and used directly in an `iSEE()` application, while the former can only be used as a parent of a concrete subclass.
Here, we will provide a brief summary of each class along with a listing of its available slots.
Readers should refer to the documentation for each class (links below) for more details.

## Virtual classes

### `Panel` 

The `r class_link("Panel")` class is the base class for all `r Biocpkg("iSEE")` panels.
It provides functionality to control general panel parameters such as the panel width and height.
It also controls the transmission of multiple row/column selections across panels.

```{r}
slotNames(getClass("Panel"))
```

### `DotPlot`

The `r class_link("DotPlot")` class inherits from the `Panel` class and is the base class for dot-based plots.
This refers to all plots where each row or column is represented by no more than one dot/point on the plot.
It provides functionality to create the plot, control the aesthetics of the points and to manage the brush/lasso selection.

```{r}
slotNames(getClass("DotPlot"))
```

### `ColumnDotPlot`

The `r class_link("ColumnDotPlot")` class inherits from the `DotPlot` class and represents all per-column dot plots.
This refers to all plots where each column is represented by no more than one dot/point on the plot.
It provides functionality to manage `colData` fields for modifying the plot aesthetics.
It is also restricted to receiving and transmitting column identities in single and multiple selections.

```{r}
slotNames(getClass("ColumnDotPlot"))
```

### `RowDotPlot`

The `r class_link("RowDotPlot")` class inherits from the `DotPlot` class and represents all per-row dot plots.
This refers to all plots where each row is represented by no more than one dot/point on the plot.
It provides functionality to manage `rowData` fields for modifying the plot aesthetics.
It is also restricted to receiving and transmitting row identities in single and multiple selections.

```{r}
slotNames(getClass("RowDotPlot"))
```

### `Table`

The `r class_link("Table")` class inherits from the `Panel` class and represents all tables rendered using `r CRANpkg("DT")`.
Each row of the table is expected to correspond to a row or column of the `SummarizedExperiment`.
This class provides functionality to render the `DT::datatable` widget, monitor single selections and apply search filters.

```{r}
slotNames(getClass("Table"))
```

### `ColumnTable`

The `r class_link("ColumnTable")` class inherits from the `Table` class and represents all tables where the rows correspond to columns of the `SummarizedExperiment`.
Instances of this class can only transmit single and multiple selections on columns.

```{r}
slotNames(getClass("ColumnTable"))
```

### `RowTable`

The `r class_link("RowTable")` class inherits from the `Table` class and represents all tables where the rows correspond to rows of the `SummarizedExperiment`.
Instances of this class can only transmit single and multiple selections on rows.

```{r}
slotNames(getClass("RowTable"))
```

## Concrete classes

### `ReducedDimensionPlot`

The `r class_link("ReducedDimensionPlot")` class inherits from the `ColumnDotPlot` class and plots reduced dimension coordinates from an entry of the `reducedDims` in a `SingleCellExperiment`.
It provides functionality to choose the result and extract the relevant entry in preparation for plotting.

```{r}
slotNames(getClass("ReducedDimensionPlot"))
```

### `FeatureAssayPlot`

The `r class_link("FeatureAssayPlot")` class inherits from the `ColumnDotPlot` class and plots the assay values for a feature across all samples, using an entry of the `assays()` from any `SummarizedExperiment` object.
It provides functionality to choose the feature of interest and any associated variable to plot on the x-axis, as well as a method to extract the relevant pieces of data in preparation for plotting.

```{r}
slotNames(getClass("FeatureAssayPlot"))
```

### `ColumnDataPlot`

The `r class_link("ColumnDataPlot")` class inherits from the `ColumnDotPlot` class and plots `colData` variables by themselves or against each other.
It provides functionality to choose and extract the variables to plot.

```{r}
slotNames(getClass("ColumnDataPlot"))
```

### `SampleAssayPlot`

The `r class_link("SampleAssayPlot")` class inherits from the `RowDotPlot` class and plots the assay values for a sample across all features, using an entry of the `assays()` from any `SummarizedExperiment` object.
It provides functionality to choose the sample of interest and any associated variable to plot on the x-axis, as well as a method to extract the relevant pieces of data in preparation for plotting.

```{r}
slotNames(getClass("FeatureAssayPlot"))
```

### `RowDataPlot`

The `r class_link("RowDataPlot")` class inherits from the `RowDotPlot` class and plots `rowData` variables by themselves or against each other.
It provides functionality to choose and extract the variables to plot.

```{r}
slotNames(getClass("RowDataPlot"))
```

### `ColumnDataTable`

The `r class_link("ColumnDataTable")` class inherits from the `ColumnTable` class and shows the contents of the `colData` in a table.
It provides functionality to extract the `colData` in preparation for rendering.

```{r}
slotNames(getClass("ColumnDataTable"))
```

### `RowDataTable`

The `r class_link("RowDataTable")` class inherits from the `RowTable` class and shows the contents of the `rowData` in a table.
It provides functionality to extract the `rowData` in preparation for rendering.

```{r}
slotNames(getClass("RowDataTable"))
```

### `ComplexHeatmapPlot`

The `r class_link("ComplexHeatmapPlot")` class inherits from the `Panel` class and creates a heatmap from assay values using the `r Biocpkg("ComplexHeatmap")` package.
It provides functionality to specify the features to be shown, which assay to show, transformations to be applied, and which metadata variables to display as row and column heatmap annotations.

```{r}
slotNames(getClass("ComplexHeatmapPlot"))
```
