--- 
title: "Extending iSEE"
author: Kevin Rue-Albrecht, Federico Marini, Charlotte Soneson, and Aaron Lun
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
description: "This is a minimal example of using the bookdown package to write a book. The output format for this example is bookdown::gitbook."
---

```{r, include=FALSE}
library(BiocStyle)
```

# Preface {-}

The [Bioconductor](https://bioconductor.org/) `r Biocpkg("iSEE")` package provides functions for creating an interactive graphical user interface (GUI) using the [RStudio])(https://rstudio.com/) `r CRANpkg("Shiny")` package for exploring data stored in SummarizedExperiment objects, including row- and column-level metadata.
In this book we describe how to create web-applications that leverage the built-in panels and develop new ones.

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```
