--- 
title: "Extending iSEE"
author: Kevin Rue-Albrecht, Federico Marini, Charlotte Soneson, and Aaron Lun
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
description: "This is a minimal example of using the bookdown package to write a book. The output format for this example is bookdown::gitbook."
---

```{r, include=FALSE}
library(BiocStyle)
```

# Preface {-}

The [Bioconductor](https://bioconductor.org/) `r Biocpkg("iSEE")` package provides functions for creating an interactive graphical user interface (GUI) using the [RStudio](https://rstudio.com/) `r CRANpkg("Shiny")` package for exploring data stored in SummarizedExperiment objects, including row- and column-level metadata.
In this book we describe how to create web-applications that leverage the built-in panels and develop new ones.

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```

<!--chapter:end:index.Rmd-->

# The iSEE server {#server}

Panels are defined as a hierarchy of S4 classes.

<!-- Source: https://www.overleaf.com/project/5de43f2303966900014b99c2 -->
\begin{tikzpicture}
  \node {Family: DotPlot}
    child { node {Family: ColumnDotPlot}
        child { node {RedDimPlot}}
        child { node {ColDataPlot}}
        child { node {FeatAssayPlot}}
    }
    child { node {Family: ColumnDotPlot}
        child { node {RowDataPlot}}
        child { node {SampAssayPlot}}
    }};
\end{tikzpicture}

The app memory is a list of instances 

<!--chapter:end:01-server.Rmd-->

`r if (knitr::is_html_output()) '
# References {-}
'`

<!--chapter:end:06-references.Rmd-->

